---
title: "Proposal title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
if (!require("pacman"))
  install.packages("pacman")
pacman::p_load(tidyverse, tidytuesdayR)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
Weather_Forecast_Accuracy <- tidytuesdayR::tt_load('2022-12-20')
Weather_Forecast_Accuracy <- tidytuesdayR::tt_load(2022, week = 51)

weather_forecasts <- Weather_Forecast_Accuracy$weather_forecasts
cities <- Weather_Forecast_Accuracy$cities
outlook_meanings <- Weather_Forecast_Accuracy$outlook_meanings
```

This dataset represents to Weather Forecast Accuracy and is sourced from the USA National Weather Service. It encompasses 16 months of weather forecasts and subsequent observations for 167 cities across the United States. Weather_forecast Accuracy is divided into three dimensions, `weather_forecasts`, `cities`, and `outlook_meanings`.

The first dataset we'll look at is `weather_forecasts`, which contains 651,968 observations of 10 variables. Specifically, this includes forecasted and observed temperatures, time differences between observations and forecasts, precipitation, forecast outlook and the presence of potential error. The second dataset we'll look at is `cities`, which contains 236 observations of 11 variables with geographic information such as latitude, longitude, and climate for 236 cities, including the 167 cities we'll be looking at. The last dataset, `outlook_meanings`, contains 23 observations of 2 variables and is the one that describes the aforementioned variable, `forecast_outlook`.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
